### 2024-11-12 _IDS702Projectproposal_template.qmd
adding files

---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
```
# Dataset 1 (top choice)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-1}

```

# Dataset 2

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-2}

```

#

# Dataset 3 (optional)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-3}

```

#

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

####