### 2024-11-15 _IDS702Projectproposal_template.qmd

---
title: "Project Proposal"
author: "ALL group member names here"
subtitle: "Due November 17 at 11:59pm"
format: pdf
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
```

# Dataset 1 (top choice)

**Data source:**
County Health Rankings & Roadmaps (NC 2024 County Data: https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)

**Brief description:**
This dataset contains health factor and health outcome data for each county in NC in 2024. It is provided as part of a program by the University of Wisconsin Population Health Institute () and combines data from other sources, such as the Behavioral Risk Factor Surveillance System & National Center for Health Statistics.

**Research question 1:**
What is the relationship between residential segregation (continuous variable) and juvenile arrests (continuous variable)? How is this relationship affected by the county's median household income (categorical variable, where 0=below national average, 1=above national average)?

Residential segregation = Index of dissimilarity where higher values indicate greater residential segregation between Black and white county residents (2018-2022 data).

Juvenile arrests = Rate of delinquency cases per 1000 juveniles (2021 data).

NY = 52
TX = 254
NC = 100
MN = 87
CO = 64
WA = 39
AL = 67

**Research question 2:**

-   How does school segregation level (continuous predictor) relate to school funding adequacy (categorical outcome)?

Qs:
-Is it okay if the variables' data come from different years? 
state in report and consider it a limitation
-Is it better to combine multiple years of data for the same set of counties or include 5 different states' counties (but same year)?
maybe look at year as interaction
-Is it okay if original dataset only has numerical data and we turn some of the variables into categorical?

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}
texas_full <- read.csv('https://github.com/rrandev03/Practical-Data-Science/raw/refs/heads/main/Texas.csv', skip=1)
#colnames(texas_full)
```

```{r}
texas <- texas_full[c('FIPS', 'State', 'County', "Segregation.Index", "Spending.per.Pupil", "School.Funding.Adequacy", "Segregation.Index.1", "Juvenile.Arrest.Rate")]
head(texas)
```
```{r}
texas <- texas |> 
  rename(
    Residential.Segregation.Index = Segregation.Index.1)

head(texas)
```

**Exploratory Plots:**

```{r plots-1}

```

# Dataset 2

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-2}

```

#

# Dataset 3 (optional)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of variable):

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-3}

```

#

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?
We will meet virtually on Tuesday between 8 and 9 pm.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?
If a team member will be missing a meeting, they should provide around 10-12 hours of advance notice (the morning of).

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?
We will be using WhatsApp primarily for communication, with Slack as our secondary chat. We will send Zoom invites via email to reserve our weekly meeting time (Tuesdays 8-9 pm). Ideally, team members should respond to or acknowledge messages within 1-3 hours on weekdays and 12 hours on weekends/during the night.

####