### 2024-11-15 CURJ_IDS702Projectproposal.qmd
EDA for Q2 county data

---
title: "Project Proposal"
author: "ALL group member names here"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Dataset 1 (top choice)

**Data source:** County Health Rankings & Roadmaps (NC 2024 County Data:
https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)

**Brief description:** This dataset contains health factor and health
outcome data for each county in NC in 2024. It is provided as part of a
program by the University of Wisconsin Population Health Institute ()
and combines data from other sources, such as the Behavioral Risk Factor
Surveillance System & National Center for Health Statistics.

**Research question 1:** What is the relationship between residential
segregation (continuous variable) and juvenile arrests (continuous
variable)? How is this relationship affected by the region of the US
that the county / state is in (categorical variable called Region)?

Residential segregation = Index of dissimilarity where higher values
indicate greater residential segregation between Black and white county
residents (2018-2022 data).

Juvenile arrests = Rate of delinquency cases per 1000 juveniles (2021
data).

NY = 52 TX = 254 NC = 100 MN = 87 CO = 64 WA = 39 AL = 67

**Research question 2:**

-   How does school segregation level (continuous predictor) relate to
    school funding adequacy (categorical outcome)?

Qs: -Is it okay if the variables' data come from different years? state
in report and consider it a limitation -Is it better to combine multiple
years of data for the same set of counties or include 5 different
states' counties (but same year)? maybe look at year as interaction -Is
it okay if original dataset only has numerical data and we turn some of
the variables into categorical?

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}
alabama_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Alabama.csv', skip=1)
texas_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Texas.csv', skip=1)
nc_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/North%20Carolina.csv', skip=1)
colorado_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Colorado.csv', skip=1)
washington_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Washington.csv', skip=1)
ny_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/New%20York.csv', skip=1)
minnesota_full <-  read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Minnesota.csv', skip=1)
#colnames(texas_full)
```

```{r}
desired_cols <- c('FIPS', 'State', 'County', 
"Segregation.Index",
"Spending.per.Pupil", "School.Funding.Adequacy", 
"Segregation.Index.1", "Juvenile.Arrest.Rate")

alabama <- alabama_full[desired_cols]
texas <- texas_full[desired_cols]
nc <- nc_full[desired_cols]
ny <- ny_full[desired_cols]
washington <- washington_full[desired_cols]
colorado <- colorado_full[desired_cols]
minnesota <- minnesota_full[desired_cols]

```

```{r}
texas <- texas |> 
  rename(
    School.Segregation.Index = Segregation.Index,
    Residential.Segregation.Index = Segregation.Index.1)

alabama <- alabama |> 
  rename(
    School.Segregation.Index = Segregation.Index,
    Residential.Segregation.Index = Segregation.Index.1)

nc <- nc |> 
  rename(
    School.Segregation.Index = Segregation.Index,
    Residential.Segregation.Index = Segregation.Index.1)

ny <- ny |> 
  rename(
    School.Segregation.Index = Segregation.Index,
    Residential.Segregation.Index = Segregation.Index.1)

washington <- washington |> 
  rename(
    School.Segregation.Index = Segregation.Index,
    Residential.Segregation.Index = Segregation.Index.1)

colorado <- colorado |> 
  rename(
    School.Segregation.Index = Segregation.Index,
    Residential.Segregation.Index = Segregation.Index.1)

minnesota <- minnesota |> 
  rename(
    School.Segregation.Index = Segregation.Index,
    Residential.Segregation.Index = Segregation.Index.1)

#head(minnesota)
```

```{r}
all_states <- rbind(alabama, texas, nc, ny, washington, 
                    colorado, minnesota)
all_states[sample(nrow(all_states), 10), ]
```

```{r}
all_states_final <- all_states |>
  mutate(Region = factor(case_when(
    State == 'Texas' ~ 'West South Central', 
    State == 'Alabama' ~ 'East South Central',
    State == 'North Carolina' ~ 'South Atlantic',
    State == 'New York' ~ 'Middle Atlantic',
    State == 'Washington' ~ 'Pacific',
    State == 'Colorado' ~ 'Mountain',
    State == 'Minnesota' ~ 'West North Central',
    TRUE ~ State #this says to keep all other values the same
  )), School.Funding.Cat = factor(case_when(
    School.Funding.Adequacy < 0 ~ 'Inadequate', 
    School.Funding.Adequacy > 0  ~ 'Adequate',
  TRUE ~ NA
  ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))

all_states_final[sample(nrow(all_states_final), 10), 
c('State', 'County', 'School.Funding.Adequacy', 'Region', 
  'School.Funding.Cat')]
```

**Exploratory Plots:** 
Q1 Outcome Variable (Juvenile Arrests):
```{r plots-1}

```

Q1 Relationship of Interest (Residential Segregation vs. Juvenile Arrests):
```{r plots-2}

```

Q2 Outcome Variable (School Funding Adequacy):

```{r}
ggplot(all_states_final, aes(x = State, fill=School.Funding.Cat)) + 
  geom_bar() + labs(x="State", y="Count of Counties", 
  fill='School Funding Adequacy', title = "School Funding Adequacy 
  in Seven US States") + theme(axis.text.x = 
  element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_discrete(labels=c('Inadequate Funding', 
  'Adequate Funding', 'No Data Available'))
```

Q2 Relationship of Interest (School Segregation & School Funding
Adequacy):

```{r plots-4}
ggplot(all_states_final, aes(x= School.Segregation.Index, y= School.Funding.Cat))+
  geom_jitter()+ 
  labs(title = "School Segregation vs. School Funding Adequacy 
       in Counties Across 7 US States",
       x = "School Segregation Index",
       y = "School Funding Adequacy") 
```

# Dataset 2

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of
    variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of
    variable):

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-5}

```

#

# Dataset 3 (optional)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of
    variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of
    variable):

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-3}

```

#

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

We will meet virtually on Tuesday between 8 and 9 pm.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

If a team member will be missing a meeting, they should provide around
10-12 hours of advance notice (the morning of).

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

We will be using WhatsApp primarily for communication, with Slack as our
secondary chat. We will send Zoom invites via email to reserve our
weekly meeting time (Tuesdays 8-9 pm). Ideally, team members should
respond to or acknowledge messages within 1-3 hours on weekdays and 12
hours on weekends/during the night.

####