### 2024-11-17 CURJ_IDS702Projectproposal.qmd
Completed EDA for county

---
title: "Project Proposal"
author: "ALL group member names here"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(GGally)
library(patchwork)
```

# Dataset 1 (top choice)

**Data source:** County Health Rankings & Roadmaps
https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)

Secondary dataset: https://data.census.gov/table/ACSDT1Y2018.B25103?q=property

**Brief description:** The source of the datasets includes health
factor and health outcome data for each county in the US by state in
2024. These datasets are provided as part of a program by the University
of Wisconsin Population Health Institute and combines data from other
sources, such as the Behavioral Risk Factor Surveillance System &
National Center for Health Statistics. We are compiling the datasets for
seven states in seven different regions of the US to create one dataset
(North Carolina, Texas, Alabama, Colorado, Minnesota, Washington, & New
York). We will also be merging in a second American Community Survey dataset to obtain property tax data for our second research question.

**Research question 1:** What is the relationship between residential
segregation (continuous variable), percentage of the county population
with limited access to healthy foods (continuous variable), and
percentage of adults who are uninsured (continuous variable), and
average life expectancy (continuous variable)? How is the relationship
between residential segregation and life expectancy affected by the
region of the US that the county / state is in (categorical variable
that we will create called Region)? The outcome variable is average life
expectancy. The predictors are residential segregation, percentage with
limited access to healthy foods, and percentage of adults who are
uninsured.

Average life expectancy = average number of years a person is expected
to live (data from National Center for Health Statistics - Natality and
Mortality Files; Census Population Estimates Program, 2019-2021).

Residential segregation = Index of dissimilarity where higher values
indicate greater residential segregation between Black and white county
residents (data from American Community Survey 5-year estimate,
2018-2022).

Percentage limited access to healthy foods = Percentage of population
who are low-income and do not live close to a grocery store (data from
USDA Food Environment Atlas, 2019).

Percentage adults uninsured = Percentage of adults under age 65 without
health insurance (data from Small Area Health Insurance Estimates,
2021).

**Research question 2:** What is the relationship between residential
segregation (continuous variable), property tax (continuous variable),
and median household income (continuous variable), on school funding
adequacy (categorical variable that we will create)?

The outcome variable is school funding adequacy, with 0=inadequate
funding and 1=adequate funding. The predictors are residential
segregation, property tax, and median household income.

School funding adequacy = The average gap in dollars between actual and
required spending per pupil among public school districts. Required
spending is an estimate of dollars needed to achieve U.S. average test
scores in each district. (data from School Finance Indicators Database,
2021).

Median property tax = Median property tax value for each county between
2018-2022 (data from IPUMS).

Median household income = The income where half of households in a
county earn more and half of households earn less (data from Small Area
Income and Poverty Estimates; American Community Survey, 5-year
estimates, 2022 & 2018-2022).

**Load the datasets and prepare final dataset`**:

```{r data-1, warning=FALSE}
alabama_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Alabama.csv', skip=1)
texas_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Texas.csv', skip=1)
nc_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/North%20Carolina.csv', skip=1)
colorado_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Colorado.csv', skip=1)
washington_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Washington.csv', skip=1)
ny_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/New%20York.csv', skip=1)
minnesota_full <-  read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Minnesota.csv', skip=1)
#colnames(alabama_full)
```

```{r}
desired_cols <- c('FIPS', 'State', 'County', 
"Life.Expectancy", 
"Segregation.Index.1", "X..Limited.Access.to.Healthy.Foods.1", 
"X..Uninsured.Adults.1", 
"Spending.per.Pupil", "School.Funding.Adequacy", 
"Median.Household.Income" 
)

alabama <- alabama_full[desired_cols]
texas <- texas_full[desired_cols]
nc <- nc_full[desired_cols]
ny <- ny_full[desired_cols]
washington <- washington_full[desired_cols]
colorado <- colorado_full[desired_cols]
minnesota <- minnesota_full[desired_cols]
#head(alabama)
```

```{r}

texas <- texas |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

alabama <- alabama |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

nc <- nc |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

ny <- ny |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

washington <- washington |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

colorado <- colorado |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

minnesota <- minnesota |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

#head(minnesota)
```

```{r}
#Remove the first row from each dataset before concatenating 
#because it contains statewide data instead of a specific county
all_states <- rbind(alabama[-1,], texas[-1,], nc[-1,], ny[-1,], 
                    washington[-1,], colorado[-1,], minnesota[-1,]) 
#all_states[sample(nrow(all_states), 10), ]
```

```{r}
all_states_final <- all_states |>
  mutate(
    Region = factor(case_when(
    State == 'Texas' ~ 'West South Central', 
    State == 'Alabama' ~ 'East South Central',
    State == 'North Carolina' ~ 'South Atlantic',
    State == 'New York' ~ 'Middle Atlantic',
    State == 'Washington' ~ 'Pacific',
    State == 'Colorado' ~ 'Mountain',
    State == 'Minnesota' ~ 'West North Central',
    TRUE ~ State #this says to keep all other values the same
  )), School.Funding.Cat = factor(case_when(
    School.Funding.Adequacy < 0 ~ 'Inadequate', 
    School.Funding.Adequacy > 0  ~ 'Adequate',
  TRUE ~ NA
  ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))

```

```{r}
prop_tax <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/acs_proptax.csv')
```

```{r}
merged_df <- merge(all_states_final, prop_tax, by = "FIPS", 
              all.x = TRUE, all.y = TRUE)

#merged_df[sample(nrow(all_states_final), 10), ]
#colSums(is.na(final_df))

```

```{r}
final_df <- merged_df |> 
  mutate(
    Median.Prop.Tax = as.numeric(med_re_.taxes_tot))
final_df = subset(final_df, select = -c(med_re_.taxes_tot, geo_id, county_name, state_name) )

```

**Provide a `glimpse():**

```{r}
glimpse(final_df)
```

**Exploratory Plots:** Q1 Outcome Variable (Life Expectancy):

```{r plots-1, warning=FALSE}
ggplot(final_df, aes(x = Life.Expectancy)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue") +
  facet_wrap(~ Region) +
  labs(title = "Life Expectancy Distribution", x = "Life Expectancy", y = "Count")
```

Q1 Relationship of Interest:

```{r plots-2, warning=FALSE}
plot1 <- ggplot(final_df, aes(x = Residential.Segregation.Index, y = Life.Expectancy)) +
  geom_point(color='purple') +
  labs(x = "Residential Segregation", y = "Life Expectancy")+ 
  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
axis.title.x = element_text(size = 8),
axis.title.y = element_text(size = 8))

plot2 <- ggplot(final_df, aes(x = Percent.Limited.Access.Healthy.Foods
, y = Life.Expectancy)) +
  geom_point(color = "green") +
  labs( x = "% Population With Limited Healthy Food Access") +
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot3 <- ggplot(final_df, aes(x = Percent.Uninsured.Adults, y = Life.Expectancy)) +
  geom_point(color = "blue") +
  labs(x = "% Adult Population Without Health Insurance", y = "Life Expectancy")+ 
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot1 + plot2 + plot3 + plot_layout(ncol = 2, nrow=2) + 
  plot_annotation(title = "Relationships Between Predictors and 
                  Life Expectancy Outcome",
                  theme = theme(
      plot.title = element_text(hjust=0.5)))
```

Q2 Outcome Variable (School Funding Adequacy):

```{r}
ggplot(final_df, aes(x = State, fill=School.Funding.Cat)) + 
  geom_bar() + labs(x="State", y="Count of Counties", 
  fill='School Funding Adequacy', title = "School Funding Adequacy 
  in Seven US States") + theme(axis.text.x = 
  element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_discrete(labels=c('Inadequate Funding', 
  'Adequate Funding', 'No Data Available'))
```

Q2 Relationship of Interest:

```{r plots-4, warning=FALSE}
ggpairs(final_df, columns = c("Residential.Segregation.Index", 
                              "Median.Household.Income",
                              "Median.Prop.Tax"), 
        aes(color = School.Funding.Cat)) + 
        labs(title = "Pairwise Plots of Predictors Colored by 
             School Funding Adequacy",
              x = "Predictor 1",
              y = "Predictor 2",
              color = "School Funding Adequacy") +
         theme(plot.title = element_text(size=15, hjust=0.5), 
               strip.text.x = element_text(size = 5),
               strip.text.y = element_text(size = 4))

```

The above figure is a pairwise scatterplot for each pair of predictor
variables (residential segregation, property tax, and household income),
with the points being colored by the school funding adequacy outcome.
Along the diagonal are histograms showing the distribution of each
individual predictor, separated by each level of the school funding
adequacy outcome variable (red = 0, or inadequate funding; blue = 1, or
adequate funding; gray = NA values). The upper half shows the
correlation coefficient between each pair of predictors, and also the
specific correlation coefficients when the data is separated out by each
level of the outcome.

# Dataset 2

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of
    variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of
    variable):

**Load the data and provide a `glimpse()`**:

```{r data-2, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-5}

```

#

# Dataset 3 (optional)

**Data source:**

**Brief description:**

**Research question 1:**

-   Outcome variable (include the name/description and type of
    variable):

**Research question 2:**

-   Outcome variable (include the name/description and type of
    variable):

**Load the data and provide a `glimpse()`**:

```{r data-3, warning=FALSE}

```

**Exploratory Plots:**

```{r plots-3}

```

#

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

We will meet virtually on Tuesday between 8 and 9 pm.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

If a team member will be missing a meeting, they should provide around
10-12 hours of advance notice (the morning of).

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

We will be using WhatsApp primarily for communication, with Slack as our
secondary chat. We will send Zoom invites via email to reserve our
weekly meeting time (Tuesdays 8-9 pm). Ideally, team members should
respond to or acknowledge messages within 1-3 hours on weekdays and 12
hours on weekends/during the night.

####