### 2024-11-17 NFL_IDS702Projectproposal.qmd
NFL data set question 2 EDA

---
title: "Project Proposal"
author: "ALL group member names here"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(dplyr)
```

# NFL Dataset

**Data source:** NFL Kaggle 2025 Play Dataset
https://www.kaggle.com/competitions/nfl-big-data-bowl-2025/data?select=player_play.csv

**Brief description:** The Play Data Dataset provides detailed
play-by-play information about football games, capturing various aspects
of gameplay, team dynamics, and player actions. This dataset includes
variables that tell us more about what influences game play within a
football game.

**Research question 1:** Question: Does the distance that a quarterback
drops back after the snap affect the probability of a pass completion?
By focusing on quarterback actions and pass outcomes, it could help
understand if dropping back type was effective or not.

Outcome Variable: Pass Result (passResult) - This is a binary variable,
with categories such as "Complete" (C) and "Incomplete" (I), indicating
whether the pass was successfully completed.

Independent Variables: **dropbackType**: The type of drop back after the
snap by the QB (Traditional, Designed Rollout, Scramble, Scramble
Rollout, Designed Rollout Left, Designed Rollout Right, Scramble Rollout
Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
(categorical)\
**dropbackDistance**: The distance the QB dropped back (yards) behind
the center after the snap (numeric)\
**timeToSack**: Time before the QB is sacked; this indicates the level
of pass protection (numeric) **rushLocationType**: Indicates where
defensive pressure might be coming from; outside pressure can force
quicker throws. (categorical)

Possible other variables:\
**pff_passCoverage**: Defensive coverage scheme (e.g., man vs. zone);
some schemes are harder to throw against.(categorica)\
**pff_manZone**: Specific indicator of man or zone coverage.
(categorical)\
**unblockedPressure**: Whether there was pressure from an unblocked
defender; unblocked pressure typically lowers completion probability.
(binary)\
**passTippedAtLine**: Whether the pass was tipped; this directly impacts
completion probability.(binary)

**Research question 2:** Question: How does the game situation affect
win probability change?

Outcome Variable: The outcome variable would be the
homeTeamWinProbabilityAdded

Independent Variable: **quarter**: the quarter the play is in may impact
the delta of the teams probability of winning as any slip up in the
latter half of the game may be more determental and harder to come back
from.\
**down**: how many downs are left in the play impacts when the team has
the ball\
**yardsToGo:** how many yards they have left to increase scoring
opportunities\
**gameClock:** how much time they have left to increase scoring
opportunities\
**preSnapHomeScore**: how far behind the are may impact their likelihood
of coming back or their hold on the game\
**preSnapVisitorScore**: how far ahead the opponent may be may impact
their likelihood of the team's coming back or hold on the game\
\
We can add an interaction term for quarter and down as that will test
whether the effect of down and distance on win probability change is
stronger or weaker in certain quarters.

**Load the data and provide a `glimpse()`**:

```{r data-1, warning=FALSE}
nfldata <- read.csv("https://raw.githubusercontent.com/wujenny214/F24_STATS_FINAL/refs/heads/main/data/NFL/plays.csv") #replace with you file path name 
glimpse(nfldata) 
```

**Exploratory Plots:**\
\
RQ1: Does the distance that a quarterback drops back after the snap
affect the probability of a pass completion?

```{r}
#Build a histogram of the win probably delta, "homeTeamWinProbabilityAdded" 
library(ggplot2)

ggplot(nfldata, aes(x = as.factor(quarter), y = preSnapHomeTeamWinProbability)) +
  geom_histogram() +
  labs(title = "Win Probability Before Play by Quarter (Home Team)", x = "Quarter", y = "Home Team Win Probability")
```

RQ2: How does the game situation affect win probability change?

```{r}
#Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. maybe a graph on gameClock and homeTeamWinProbabilityAdded
ggplot(nfldata, aes(x = homeTeamWinProbabilityAdded)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue") +  # Adjust binwidth as needed
  labs(title = "Home Team Win Probability Added Distribution",
       x = "Home Team Win Probability Added",
       y = "Count") +
  theme_minimal()  # Minimal theme for a clean look
```

```{r}
#Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. maybe a graph on gameClock and homeTeamWinProbabilityAdded
# Load the necessary packages
library(ggplot2)
library(patchwork)

# Create the individual plots
plot1 <- ggplot(nfldata, aes(x = factor(down), y = homeTeamWinProbabilityAdded)) +
  geom_boxplot(fill = 'purple') +
  labs(x = "Down", y = "Home Team Win Probability Added") + 
  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot2 <- ggplot(nfldata, aes(x = factor(quarter), y = homeTeamWinProbabilityAdded)) +
  geom_boxplot(fill = "green") +
  labs(x = "Quarter", y = "Win Probability") +
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot3 <- ggplot(nfldata, aes(x = preSnapVisitorScore, y = homeTeamWinProbabilityAdded)) +
  geom_point(color = "red") +
  labs(x = "Pre-Snap Visitor Score", y = "Win Probability") +
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

# Combine the plots using patchwork
combined_plot <- plot1 + plot2 + plot3 + 
  plot_layout(ncol = 2, nrow = 2) + 
  plot_annotation(title = "Relationships Between Predictors and Win Probability Outcome",
                  theme = theme(plot.title = element_text(hjust = 0.5)))

# Print the combined plot
print(combined_plot)

```