### 2024-11-27 project_proposal/Final_Proposal_CURJ_IDS702.qmd
deleted nfl data and added a project_proposal folder

---
title: "Project Proposal"
author: "Jenny Wu, Rishika Randev, Shiyue Zhou, Uzoma Uwazurike"
subtitle: "Due November 17 at 11:59pm"
format: pdf
editor: 
  markdown: 
    wrap: 72
echo: FALSE
warning: FALSE
---

#### Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(GGally)
library(patchwork)
```

# Dataset 1 (top choice)

**Data source:** County Health Rankings & Roadmaps
https://www.countyhealthrankings.org/health-data/north-carolina/data-and-resources)

Secondary dataset: https://data.census.gov/table/ACSDT1Y2018.B25103?q=property

**Brief description:** The source of the datasets includes health
factor and health outcome data for each county in the US by state in
2024. These datasets are provided as part of a program by the University
of Wisconsin Population Health Institute and combines data from other
sources, such as the Behavioral Risk Factor Surveillance System &
National Center for Health Statistics. We are compiling the datasets for
seven states in seven different regions of the US to create one dataset
(North Carolina, Texas, Alabama, Colorado, Minnesota, Washington, & New
York). We will also be merging in a second American Community Survey dataset to obtain property tax data for our second research question.

**Research question 1:** What is the relationship between residential
segregation (continuous variable), percentage of the county population
with limited access to healthy foods (continuous variable), and
percentage of adults who are uninsured (continuous variable), and
average life expectancy (continuous variable)? How is the relationship
between residential segregation and life expectancy affected by the
region of the US that the county / state is in (categorical variable
that we will create called Region)? The outcome variable is average life
expectancy. The predictors are residential segregation, percentage with
limited access to healthy foods, and percentage of adults who are
uninsured.

Average life expectancy = average number of years a person is expected
to live (data from National Center for Health Statistics - Natality and
Mortality Files; Census Population Estimates Program, 2019-2021).

Residential segregation = Index of dissimilarity where higher values
indicate greater residential segregation between Black and white county
residents (data from American Community Survey 5-year estimate,
2018-2022).

Percentage limited access to healthy foods = Percentage of population
who are low-income and do not live close to a grocery store (data from
USDA Food Environment Atlas, 2019).

Percentage adults uninsured = Percentage of adults under age 65 without
health insurance (data from Small Area Health Insurance Estimates,
2021).

**Research question 2:** What is the relationship between residential
segregation (continuous variable), property tax (continuous variable),
and median household income (continuous variable), on school funding
adequacy (categorical variable that we will create)?

The outcome variable is school funding adequacy, with 0=inadequate
funding and 1=adequate funding. The predictors are residential
segregation, property tax, and median household income.

School funding adequacy = The average gap in dollars between actual and
required spending per pupil among public school districts. Required
spending is an estimate of dollars needed to achieve U.S. average test
scores in each district. (data from School Finance Indicators Database,
2021).

Median property tax = Median property tax value for each county between
2018-2022 (data from IPUMS).

Median household income = The income where half of households in a
county earn more and half of households earn less (data from Small Area
Income and Poverty Estimates; American Community Survey, 5-year
estimates, 2022 & 2018-2022).

**Load the datasets and prepare final dataset (code is hidden to reduce proposal length)**:

```{r data-1, warning=FALSE}
alabama_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Alabama.csv', skip=1)
texas_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Texas.csv', skip=1)
nc_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/North%20Carolina.csv', skip=1)
colorado_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Colorado.csv', skip=1)
washington_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Washington.csv', skip=1)
ny_full <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/New%20York.csv', skip=1)
minnesota_full <-  read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/Minnesota.csv', skip=1)
#colnames(alabama_full)
```

```{r}
desired_cols <- c('FIPS', 'State', 'County', 
"Life.Expectancy", 
"Segregation.Index.1", "X..Limited.Access.to.Healthy.Foods.1", 
"X..Uninsured.Adults.1", 
"Spending.per.Pupil", "School.Funding.Adequacy", 
"Median.Household.Income" 
)

alabama <- alabama_full[desired_cols]
texas <- texas_full[desired_cols]
nc <- nc_full[desired_cols]
ny <- ny_full[desired_cols]
washington <- washington_full[desired_cols]
colorado <- colorado_full[desired_cols]
minnesota <- minnesota_full[desired_cols]
#head(alabama)
```

```{r}

texas <- texas |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

alabama <- alabama |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

nc <- nc |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

ny <- ny |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

washington <- washington |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

colorado <- colorado |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

minnesota <- minnesota |> 
  rename(
    Percent.Uninsured.Adults = X..Uninsured.Adults.1, 
    Percent.Limited.Access.Healthy.Foods = X..Limited.Access.to.Healthy.Foods.1,
    Residential.Segregation.Index = Segregation.Index.1)

#head(minnesota)
```

```{r}
#Remove the first row from each dataset before concatenating 
#because it contains statewide data instead of a specific county
all_states <- rbind(alabama[-1,], texas[-1,], nc[-1,], ny[-1,], 
                    washington[-1,], colorado[-1,], minnesota[-1,]) 
#all_states[sample(nrow(all_states), 10), ]
```

```{r}
all_states_final <- all_states |>
  mutate(
    Region = factor(case_when(
    State == 'Texas' ~ 'West South Central', 
    State == 'Alabama' ~ 'East South Central',
    State == 'North Carolina' ~ 'South Atlantic',
    State == 'New York' ~ 'Middle Atlantic',
    State == 'Washington' ~ 'Pacific',
    State == 'Colorado' ~ 'Mountain',
    State == 'Minnesota' ~ 'West North Central',
    TRUE ~ State #this says to keep all other values the same
  )), School.Funding.Cat = factor(case_when(
    School.Funding.Adequacy < 0 ~ 'Inadequate', 
    School.Funding.Adequacy > 0  ~ 'Adequate',
  TRUE ~ NA
  ), levels=c('Inadequate', 'Adequate'), labels=c(0, 1)))

```

```{r}
prop_tax <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/county_data/acs_proptax.csv')
```

```{r}

merged_df <- merge(all_states_final, prop_tax, by = "FIPS", 
              all.x = TRUE, all.y = TRUE)

#merged_df[sample(nrow(all_states_final), 10), ]
#colSums(is.na(final_df))
```

```{r}
final_df <- merged_df |> 
  mutate(
    Median.Prop.Tax = as.numeric(med_re_.taxes_tot))
final_df = subset(final_df, select = -c(med_re_.taxes_tot, geo_id, county_name, state_name) )

```

**Provide a `glimpse():**

```{r}
glimpse(final_df)
```

**Exploratory Plots:** 
Q1 Outcome Variable (Life Expectancy):

```{r plots-1, warning=FALSE}
ggplot(final_df, aes(x = Life.Expectancy)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue") +
  facet_wrap(~ Region) +
  labs(title = "Life Expectancy Distribution", x = "Life Expectancy", y = "Count")
```

Q1 Relationship of Interest:

```{r plots-2, warning=FALSE}
plot1 <- ggplot(final_df, aes(x = Residential.Segregation.Index, y = Life.Expectancy)) +
  geom_point(color='purple') +
  labs(x = "Residential Segregation", y = "Life Expectancy")+ 
  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
axis.title.x = element_text(size = 8),
axis.title.y = element_text(size = 8))

plot2 <- ggplot(final_df, aes(x = Percent.Limited.Access.Healthy.Foods
, y = Life.Expectancy)) +
  geom_point(color = "green") +
  labs( x = "% Population With Limited Healthy Food Access") +
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot3 <- ggplot(final_df, aes(x = Percent.Uninsured.Adults, y = Life.Expectancy)) +
  geom_point(color = "blue") +
  labs(x = "% Adult Population Without Health Insurance", y = "Life Expectancy")+ 
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot1 + plot2 + plot3 + plot_layout(ncol = 2, nrow=2) + 
  plot_annotation(title = "Relationships Between Predictors and 
                  Life Expectancy Outcome",
                  theme = theme(
      plot.title = element_text(hjust=0.5)))
```

Q2 Outcome Variable (School Funding Adequacy):

```{r}
ggplot(final_df, aes(x = State, fill=School.Funding.Cat)) + 
  geom_bar() + labs(x="State", y="Count of Counties", 
  fill='School Funding Adequacy', title = "School Funding Adequacy 
  in Seven US States") + theme(axis.text.x = 
  element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_discrete(labels=c('Inadequate Funding', 
  'Adequate Funding', 'No Data Available'))
```

Q2 Relationship of Interest:

```{r plots-4, warning=FALSE}
ggpairs(final_df, columns = c("Residential.Segregation.Index", 
                              "Median.Household.Income",
                              "Median.Prop.Tax"), 
        aes(color = School.Funding.Cat)) + 
        labs(title = "Pairwise Plots of Predictors Colored by 
             School Funding Adequacy",
              x = "Predictor 1",
              y = "Predictor 2",
              color = "School Funding Adequacy") +
         theme(plot.title = element_text(size=15, hjust=0.5), 
               strip.text.x = element_text(size = 5),
               strip.text.y = element_text(size = 3))

```

The above figure is a pairwise scatterplot for each pair of predictor
variables (residential segregation, property tax, and household income),
with the points being colored by the school funding adequacy outcome.
Along the diagonal are histograms showing the distribution of each
individual predictor, separated by each level of the school funding
adequacy outcome variable (red = 0, or inadequate funding; blue = 1, or
adequate funding; gray = NA values). The upper half shows the
correlation coefficient between each pair of predictors, and also the
specific correlation coefficients when the data is separated out by each
level of the outcome.

# Dataset 2

**Data source:** NFL Kaggle 2025 Play Dataset
https://www.kaggle.com/competitions/nfl-big-data-bowl-2025/data?select=player_play.csv

**Brief description:** The Play Data Dataset provides detailed
play-by-play information about football games, capturing various aspects
of gameplay, team dynamics, and player actions. This dataset includes
variables that tell us more about what influences game play within a
football game.

**Research question 1:** Does the distance that a quarterback
drops back after the snap and the dropback type affect the probability of a pass completion?

**Outcome Variable**: `passResult` (Binary: Complete/Incomplete),
indicating whether the pass was successfully completed.

**Primary Predictors**:

-   `dropbackDistance` (Continuous).The distance the QB dropped back
    (yards) behind the center after the snap (numeric)

-   `dropbackType` (Categorical).The type of drop back after the snap by
    the QB (Traditional, Designed Rollout, Scramble, Scramble Rollout,
    Designed Rollout Left, Designed Rollout Right, Scramble Rollout
    Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
    (categorical)

**Potential Interaction Terms:**

-   `dropbackDistance* dropbackType` : Different types of dropbacks
    (e.g., "Rollout" vs. "Traditional") may influence the effectiveness
    of a specific dropback distance.

**Research question 2:** How does the game situation (down, quarter, and opponent score) affect
home team's win probability change after a play?

**Outcome Variable**: `homeTeamWinProbabilityAdded` (Continuous),
indicating how much the home team's win probability changes after a given play.

**Primary Predictors**:

-   `quarter` (Discrete), indicating whether the play is in the first, second, third, or fourth      quarter, or during overtime.

-   `down` (Discrete), indicating whether the play is on a first, second, third, or fourth down.

-   `preSnapVisitorScore` (Discrete), indicating the score of the oponent team prior to the          current play.

**Load the data and provide a `glimpse()`**:

```{r, warning=FALSE}
nfldata <- read.csv("https://raw.githubusercontent.com/wujenny214/F24_STATS_FINAL/refs/heads/main/data/NFL/plays.csv") #replace with you file path name 
glimpse(nfldata) 
```

**Exploratory Plots:**
RQ1: Does the distance that a quarterback drops back after the snap
affect the probability of a pass completion?

Q1 Outcome Plot:

```{r}
# Clean and transform dataset
nfldata_1 <- nfldata %>%
  # Select relevant variables
  select(dropbackDistance, dropbackType, timeToThrow, passLength, passResult) %>%
  # Filter for valid pass results
  filter(
    passResult %in% c("C", "I"),
    !is.na(dropbackType) & dropbackType != "UNKNOWN",
    !is.na(dropbackDistance),
    !is.na(timeToThrow),
    !is.na(passLength)
  ) %>%
  # Convert variables to factors
  mutate(
    dropbackType = as.factor(dropbackType),
    passResult = as.factor(passResult)
  )
glimpse(nfldata_1)

```

```{r}
ggplot(nfldata_1, aes(x = dropbackType, fill=passResult)) + 
  geom_bar() + labs(x="dropbackType", y="dropbackDistance", 
  fill='passResult', title = "Pass Result
  by Dropback Type") + theme(axis.text.x = 
  element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_discrete(labels=c("Complete", "Incomplete"))
```

Q1 Relationship of Interest:

```{r}
table(nfldata_1$dropbackType)
```

```{r}
nfldata_1_viz <-nfldata_1 %>%
  mutate(
    dropbackCategory = case_when(
      dropbackType == "TRADITIONAL" ~ "Traditional",
      dropbackType %in% c("DESIGNED_ROLLOUT_LEFT", "DESIGNED_ROLLOUT_RIGHT") ~ "Designed Rollout",
      dropbackType %in% c("SCRAMBLE", "SCRAMBLE_ROLLOUT_LEFT", "SCRAMBLE_ROLLOUT_RIGHT") ~ "Scramble",
      TRUE ~ NA_character_  # Handle unexpected values
    )
  )

ggpairs(
  nfldata_1_viz,
  columns = c("dropbackCategory", "dropbackDistance", "timeToThrow", "passLength"),
  aes(color = as.factor(passResult)), # Treat passResult as a factor
  upper = list(continuous =  wrap("points", size = 0.3)),
  lower = list(continuous =  wrap("points", size = 0.3))
) +
  labs(
    title = "Pairwise Plots of Predictors Colored by Pass Result",
    x = "Predictor 1",
    y = "Predictor 2",
    color = "Pass Result"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(size = 3.5),  # Adjust x-axis text size
    axis.text.y = element_text(size = 8),  # Adjust y-axis text size
    axis.ticks.x = element_blank(),       # Remove excess x-axis ticks
    axis.ticks.y = element_blank(),     # Remove excess y-axis ticks
    strip.text.y = element_text(size = 3)
  )
```

RQ2: How does the game situation affect win probability change?

Q2 Outcome Plot:

```{r}
ggplot(nfldata, aes(x = homeTeamWinProbabilityAdded)) +
  geom_histogram(binwidth = 0.05, fill = "skyblue") + 
  labs(title = "Home Team Win Probability Added Distribution",
       x = "Home Team Win Probability Added",
       y = "Count") +
  theme_minimal()  
```

Q2 Relationship of Interest:
```{r}
#Create a plot that demonstrates the relationship between the game's characteristics and the probability of winning. 
# Load the necessary packages
library(ggplot2)
library(patchwork)

# Create the individual plots
plot1 <- ggplot(nfldata, aes(x = factor(down), y = homeTeamWinProbabilityAdded)) +
  geom_boxplot() +
  labs(x = "Down", y = "Win Probability Change") + 
  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot2 <- ggplot(nfldata, aes(x = factor(quarter), y = homeTeamWinProbabilityAdded)) +
  geom_boxplot() +
  labs(x = "Quarter", y = "Win Probability Change") +
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

plot3 <- ggplot(nfldata, aes(x = preSnapVisitorScore, y = homeTeamWinProbabilityAdded)) +
  geom_point(color = "red") +
  labs(x = "Pre-Snap Visitor Score", y = "Win Probability Change") +
  theme(plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8))

# Combine the plots using patchwork
combined_plot <- plot1 + plot2 + plot3 + 
  plot_layout(ncol = 2, nrow = 2) + 
  plot_annotation(title = "Relationships Between Predictors and Home Team Win Probability Change",
                  theme = theme(plot.title = element_text(size=8, hjust = 0.5)))

# Print the combined plot
print(combined_plot)

```

# Team Charter

#### When will you meet as a team to work on the project components? Will these meetings be held in person or virtually?

We will meet virtually on Tuesday between 8 and 9 pm.

#### What is your group policy on missing team meetings (e.g., how much advance notice should be provided)?

If a team member will be missing a meeting, they should provide around
10-12 hours of advance notice (the morning of).

#### How will your team communicate (email, Slack, text messages)? What is your policy on appropriate response time (within a certain number of hours? Nights/weekends?)?

We will be using WhatsApp primarily for communication, with Slack as our
secondary chat. We will send Zoom invites via email to reserve our
weekly meeting time (Tuesdays 8-9 pm). Ideally, team members should
respond to or acknowledge messages within 1-3 hours on weekdays and 12
hours on weekends/during the night.

####