### 2024-11-30 analysis.qmd
Adding analysis code file

---
title: "analysis"
format: html
editor: visual
---

## Load Data & Create Factors

```{r}
combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
```

```{r}
head(combined)
```

```{r}
combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest'), labels=c(0, 1, 2, 3)))
```

```{r}
combined <- combined %>% select(-X, -Spending.per.Pupil)
```

```{r}
head(combined)
```

## Imputation

```{r}

library(mice)
library(sjlabelled)
library(tidyverse)

```

```{r}
sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
sub.imp <- mice(sub, m=5, method="pmm", print=FALSE)
```

```{r}
#Need to convert school funding adequacy imputed values to school funding categorical variable before using in models
```

## Q1

```{r}
#MLR
#Outcome = life expectancy
#Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)
```

## Q2

```{r}
#Logistic 
#Outcome = School.Funding.Cat
#Predictors = residential segregation, median.household.income, median.prop.tax
```