### 2024-11-30 missing.qmd

---
title: "cleaning.qmd"
format: html
editor: visual
---

```{r}
library(dplyr)
library(tidyverse)
library(Hmisc)
```

```{r}
combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
```

```{r}
combined = subset(combined, select = -c(X))
head(combined)

```

```{r}
dim(combined)
```

```{r}
combined |> group_by(State) %>% summarise(total_count = n())
```

```{r}
sum(duplicated(combined))
```

```{r}
sum(duplicated(combined$FIPS))
```

```{r}
summary(combined)
```

```{r}
combined <- combined %>%
    rename(
      "Perc.Uninsured.Adults" = "X..Uninsured.Adults",
      "Perc.Limited.Access.to.Healthy.Foods" = "X..Limited.Access.to.Healthy.Foods"
    )
```

```{r}
combined %>%
           group_by(State) %>%
           summarise(total_count = n(), Income_NA_count = sum(is.na(Median.Household.Income)), Uninsured_NA_count = sum(is.na(Perc.Uninsured.Adults)), Tax_NA_count = sum(is.na(Median.Prop.Tax)), Segregation_NA_count = sum(is.na(Residential.Segregation.Index)), Foods_NA_count = sum(is.na(Perc.Limited.Access.to.Healthy.Foods)), Expectancy_NA_count = sum(is.na(Life.Expectancy)), Adequacy_NA_count = sum(is.na(School.Funding.Adequacy)))
```

```{r}
na.patterns <- naclus(combined)
naplot(na.patterns, "na per var")
```

```{r}
plot(na.patterns)
```

```{r}
combined <- combined |>
  mutate(expectancy.miss = ifelse(is.na(Life.Expectancy),1,0))

expectancy_miss_mod <- glm(expectancy.miss ~ Residential.Segregation.Index+Percent.Limited.Access.Healthy.Foods+Percent.Uninsured.Adults, 
                    data=combined, family="binomial")

summary(expectancy_miss_mod)
```

```{r}
combined <- combined |>
  mutate(foods.miss = ifelse(is.na(Percent.Limited.Access.Healthy.Foods),1,0))

foods_miss_mod <- glm(foods.miss ~ Residential.Segregation.Index+Life.Expectancy+Percent.Uninsured.Adults, 
                    data=combined, family="binomial")

summary(foods_miss_mod)
```

```{r}
combined <- combined |>
  mutate(funding.miss = ifelse(is.na(School.Funding.Adequacy),1,0))

funding_miss_mod <- glm(funding.miss ~ Residential.Segregation.Index+Median.Prop.Tax+Median.Household.Income, 
                    data=combined, family="binomial")

summary(funding_miss_mod)
```

```{r}
colnames(combined)
```

```{r}
library(mice)
library(sjlabelled)
library(tidyverse)

sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Cat", "Median.Prop.Tax"))
sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)
sub.imp <- mice(sub, m=5, method="pmm", print=FALSE)
```

```{r}
sub.mods <- with(sub.imp, #replace with. model
)
summary(pool(sub.mods))
```