### 2024-12-09 code.qmd
code and report qmds

---
title: "Understanding the Impact of Socioeconomic and Regional Factors on Health and Educational Outcomes"
author: "Rishika Randev, Jenny Wu, Uzoma Uwazurike Jr., Shiyue Zhou"
instructor: "Andrea Lane"
institution: "Duke University, IDS 702 Data Modeling and Presentation"
format: html
editor: visual
execute:
  echo: false
  warning: false
---

### Exploratory Data Analysis & Data Preprocessing

(Rishika will add here)

### Model Fitting & Assessment

```{r, echo= FALSE, warning = FALSE, message=FALSE}
library("dplyr")
library(mice)
library(sjlabelled)
library(tidyverse)
library("modelsummary")
library(car)
```

```{r,echo= FALSE, warning = FALSE, message=FALSE}
combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')
```

```{r, warning = FALSE}
# Making the state_fac variable a factor variable for future use
combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest')))
```

```{r, warning = FALSE}
## Imputation Code 
# In this block, we are looking to use predictive mean matching to combing regression methods and hot deck imputing to predicted value closest to the predicted value of the observation with missing data; we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models

sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)

# Add School.Funding.Cat to the orginal dataset. Since we do not want to imputate on the categorical outcome variable,  this creation of a variable exists to create the proper dataframe within the imputated data 
sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)

# Perform the imputation
sub.imp <- mice(sub, m = 1, method = "pmm", print = FALSE)

```

```{r, warning = FALSE}
# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the imputed outcome variable.

# Define the categorization function
categorize_school_funding <- function(x) {
  ifelse(x > 0, 1, 0)  # Binary classification
}

# Loop over the imputations
for (i in 1:1) {
  # Extract the imputed values for the i-th dataset
  imputed_values <- sub.imp$imp$School.Funding.Adequacy[, i]

  # Categorize the imputed values
  categorized_values <- categorize_school_funding(imputed_values)
  complete_cat <- rep(NA, nrow(sub))  # Initialize with NAs

  # Get the indices where imputation occurred
  imp_indices <- which(!is.na(categorized_values))  # Only for imputed values

  # Assign the categorized values to the rows that were imputed
  complete_cat[imp_indices] <- categorized_values[imp_indices]

  # Save into the `imp` list of the `sub.imp` object
  sub.imp$imp$School.Funding.Cat[[i]] <- complete_cat[imp_indices]
}
```

```{r, warning = FALSE}
# To have a full dataset, we are taking the first imputed dataset R provided and using that as our full dataset sample. 
completed_data <- complete(sub.imp, 1)

# Check the structure
str(completed_data)

# Ensure that there are no missing data 
colSums(is.na(completed_data))

```

## Q1 Life Expectancy; MLR

### **Research Question 1: Life Expectancy**

### Model Fitting (Original Data)

```{r, warning = FALSE}
# MLR
# Outcome = life expectancy 
# Predictors = residential segregation, limited access to healthy foods, uninsured adults, region_fac as interaction with residential segregation (or any other variable)

# To get a general sense of what the data tells us, we use the orignial dataset with missing values.

mlrmodel <- lm(
  Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
  X..Limited.Access.to.Healthy.Foods +
  X..Uninsured.Adults,
  data = combined)

# Renaming the output names 
renamed <- c(
  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
  "Region_facWest" = "Region.W", 
  "Region_facNortheast" = "Region.NE",
  "Region_facMidwest" = "Region.MidW")

summary(mlrmodel, fmt = fmt_significant(2),
             coef_rename = renamed, 
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA,
             )

```

### Model Fitting (Imputed Data)

test

```{r}
# Running a regression with the imputated data (completed_data) to understand the impact of missing data

mlr.imp <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
    X..Limited.Access.to.Healthy.Foods +
    X..Uninsured.Adults, data=completed_data)

renamed <- c(
  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
  "Region_facWest" = "Region.W", 
  "Region_facNortheast" = "Region.NE",
  "Region_facMidwest" = "Region.MidW"
  )

summary(mlr.imp, fmt = fmt_significant(2),
             coef_rename = renamed, 
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA,
             )
```

```{r echo= FALSE, warning = FALSE}
# Running a regression without the interaction term between residential segregation on the regions

mlr.imp.reduced <- lm(Life.Expectancy ~ Residential.Segregation.Index +
    X..Limited.Access.to.Healthy.Foods +
    X..Uninsured.Adults + 
    Region_fac, data=completed_data)

renamed <- c(
  "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
  "X..Limited.Access.to.Healthy.Foods" = "Access to Healthy Foods",
  "X..Uninsured.Adults" = "Pct. of Uninsured Adults",
  "Region_facWest" = "Region.W", 
  "Region_facNortheast" = "Region.NE",
  "Region_facMidwest" = "Region.MidW", 
  "Residential.Segregation.Index:Region_facWest" = "RSI * Region.W ", 
  "Residential.Segregation.Index:Region_facNortheast" = "RSI * Region.NE",
  "Residential.Segregation.Index:Region_facMidwest" = "RSI * Region.MidW"
  )

summary(mlr.imp.reduced, fmt = fmt_significant(2),
             coef_rename = renamed, 
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA,
             )

```

FINAL FULL MODEL: mlr.imp

FINAL REDUCED MODEL: mlr.imp.reduced (without region \* segregation interactino)

### Evaluation & Assessment of Model

```{r}
#QQPlot
qqnorm(residuals(mlr.imp))
qqline(residuals(mlr.imp), col = "red") 
```

```{r}
##Nested F-Test to see if interaction term makes a difference
anova(mlr.imp, mlr.imp.reduced)

```

```{r}
#Cooks Distance 
cooksd <- cooks.distance(mlr.imp)

# Plot Cook's Distance
plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
abline(h = 4/(nrow(data) - length(coef(mlr.imp))), col = "red")  # Threshold line

```

```{r include = FALSE}
# Identify Influential Points
influential <- which(cooksd > 4/(nrow(combined) - length(coef(mlr.imp))))
print(influential)
```

```{r}
#Variance Inflaction Factor (VIF)

vif(mlr.imp)
```

```{r}
vif(mlr.imp.reduced)
```

Please check first one!!

```{r}
#Generate 4 Diagnostic Plots at once
par(mfrow = c(2, 2), mar = c(4,4,2,2))
plot(mlr.imp)
```

## Q2 **School Funding Adequacy; GLM**

### **Research Question 2: School Funding Adequacy**

### Model Fitting

```{r}
## Checking to see if the imputated data has a large effect on the outcome

# Logistic Regression without Imputated Data
# Outcome = School.Funding.Cat
# Predictors = residential segregation, median.household.income, median.prop.tax

lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax,
              data=combined,
              family="binomial")

library("modelsummary")

#log-odds scale
summary(lmod,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA)

summary(lmod,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)

```

### Regression Results

```{r}
imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")

#log-odds scale
summary(imp_mods,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA)

summary(imp_mods,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)
```

### Evaluation and Assessment of Model

```{r}
# Creating the confusion matrix

library(caret)
library(pROC)

# set.seed(42)
# train_idx <- sample(1:nrow(long_data), size = 0.5 * nrow(long_data))
# train_data <- long_data[train_idx, ]
# test_data <- long_data[-train_idx, ]

#t_glm_model <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data = completed_data, family = binomial)

# Make predictions on the test set
test_probs <- predict(imp_mods, newdata = completed_data, type = "response")
test_pred <- ifelse(test_probs > 0.5, 1, 0)  # Classify based on threshold 0.5
```

```{r}
# Create a confusion matrix
library("caret")
conf_matrix <- confusionMatrix(factor(test_pred), factor(completed_data$School.Funding.Cat))
#print(conf_matrix)

library(pander)

# Format and display confusion matrix as a Markdown table
pander(conf_matrix, caption = "Confusion Matrix")

```

```{r}
#Creating the AUC curves
roc_result <- roc(
  completed_data$School.Funding.Cat,
  predict(imp_mods, type = "response"),
  legacy.axes = TRUE,
  plot = TRUE,
  print.thres = 0.5,
  print.auc = FALSE  # Disable automatic AUC
)

# Add AUC manually
text(
  x = 1.2,  # Adjust X-coordinate
  y = 1,  # Adjust Y-coordinate
  labels = sprintf("AUC = %.3f", auc(roc_result)),  # Add AUC value
  cex = 1,  # Label size
  col = "red"  # Label color
)
```

```{r}

```