### 2024-12-15 paper.qmd
adding conclusion para

---
title: "Understanding the Impact of Socioeconomic and Regional Factors on Health and Educational Outcomes"
author: "Rishika Randev, Jenny Wu, Uzoma Uwazurike Jr., Shiyue Zhou"
instructor: "Andrea Lane"
institution: "Duke University, IDS 702 Data Modeling and Presentation"
format: pdf
header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
fontsize: 10pt
editor: visual
execute:
  echo: false
  warning: false
---

```{r, echo= FALSE, warning = FALSE, message=FALSE}
library("dplyr") 
library(mice) 
library(sjlabelled) 
library(tidyverse) 
library("modelsummary") 
library(car)
library(GGally)
library(ggplot2)
library(patchwork)
library(caret)
library(pROC)
library(pander)
```

```{r,echo= FALSE, warning = FALSE, message=FALSE}
combined <- read.csv('https://github.com/wujenny214/F24_STATS_FINAL/raw/refs/heads/main/combined.csv')

```

```{r, echo= FALSE, warning = FALSE}
# Making the state_fac variable a factor variable for future use
combined <- combined |> mutate(State_fac = factor(combined$State), Region_fac = factor(combined$Region, levels=c('South', 'West', 'Northeast', 'Midwest')))

```

```{r, echo= FALSE, warning = FALSE}
## Imputation Code 
# In this block, we are looking to use predictive mean matching to combing regression methods and hot deck imputing to predicted value closest to the predicted value of the observation with missing data; we also needed to convert school funding adequacy imputed values to school funding categorical variable before using in models

sub <- combined |> select(c("Life.Expectancy","Residential.Segregation.Index","X..Limited.Access.to.Healthy.Foods","X..Uninsured.Adults","Median.Household.Income", "School.Funding.Adequacy", "Median.Prop.Tax", "Region_fac"))
sub <- unlabel(sub) #unlabel the data (labels cause problem for the mice function)

# Add School.Funding.Cat to the orginal dataset. Since we do not want to imputate on the categorical outcome variable,  this creation of a variable exists to create the proper dataframe within the imputated data 
sub$School.Funding.Cat <- ifelse(sub$School.Funding.Adequacy > 0, 1, 0)

# Perform the imputation
sub.imp <- mice(sub, m = 1, method = "pmm", print = FALSE, seed=3)

```

```{r, echo= FALSE, warning = FALSE}
# Since we should not imputate on the outcome, but need to imputate the School.Funding.Adequacy variable, we are feeding the School.Funding.Adequacy variable into a ifelse conditional that will replaced the imputed outcome variable.

# Define the categorization function
categorize_school_funding <- function(x) {
  ifelse(x > 0, 1, 0)  # Binary classification
}

# Loop over the imputations
for (i in 1:1) {
  # Extract the imputed values for the i-th dataset
  imputed_values <- sub.imp$imp$School.Funding.Adequacy[, i]

  # Categorize the imputed values
  categorized_values <- categorize_school_funding(imputed_values)
  complete_cat <- rep(NA, nrow(sub))  # Initialize with NAs

  # Get the indices where imputation occurred
  imp_indices <- which(!is.na(categorized_values))  # Only for imputed values

  # Assign the categorized values to the rows that were imputed
  complete_cat[imp_indices] <- categorized_values[imp_indices]

  # Save into the `imp` list of the `sub.imp` object
  sub.imp$imp$School.Funding.Cat[[i]] <- complete_cat[imp_indices]
}

```

```{r, echo= FALSE, warning = FALSE, include = FALSE }
# To have a full dataset, we are taking the first imputed dataset R provided and using that as our full dataset sample. 
completed_data <- complete(sub.imp, 1)

# Check the structure
str(completed_data)

# Ensure that there are no missing data 
colSums(is.na(completed_data))
```

## Abstract

This study examines the relationships between socioeconomic, health, and educational indicators in U.S. counties using data from the County Health Rankings & Roadmaps program and IPUMS. We analyze how factors such as residential segregation, health insurance coverage, access to healthy foods, and region impact life expectancy, and how residential segregation and region affect school funding adequacy. Our findings reveal that life expectancy is influenced by residential segregation and uninsured rates, though these effects are minor compared to the more substantial influence of household income and regional differences. Life expectancy averages are highest in the Northeast, followed by the Midwest, West, and South. Region also plays the most significant role out of all included variables in determining the likelihood of school funding adequacy, with schools in the Northeast and Midwest being far more likely to receive adequate funding than those in the South. Other factors, such as residential segregation, property taxes, and household income, show statistically significant but smaller effects. These results highlight the complex interplay between socioeconomic variables and geographic location in shaping public well-being. By identifying key disparities in health and education, this study underscores the importance of equitable resource allocation and targeted policies to address regional and socioeconomic inequalities.

## **Introduction**

Life expectancy and school funding are important indicators of public well-being because they capture fundamental aspects of health, education, and social equity that are essential for the development of individuals and communities. Life expectancy is an important indicator that reflects not only the overall health of a society, but also its stability, economic resilience, and levels of inequality. While access to quality healthcare and nutritious foods are drivers of well-being, where you live and the nature of the community you live in likely also have an influence on health outcomes such as life expectancy. According to "The Root Causes of Health Inequity - Communities in Action" report from the National Institutes of Health, we can infer that differences in life expectancy between regions and socioeconomic groups often reveal systemic inequalities, with marginalized communities experiencing shorter life spans due to limited access to health care, healthy living environments, and economic opportunities.^1^ When it comes to educational indicators, adequately funded schools ensure that students have equal access to learning opportunities, serving as a crucial investment in social mobility. The Stanford Center for Opportunity Policy in Education highlights that "the unequal allocation and inadequate levels of resources in schools and communities \[are\] at the heart of many gaps in student opportunity."^2^ This emphasizes the importance of equitable and adequate school funding in closing opportunity gaps and empowering students from disadvantaged backgrounds to achieve academic and economic success. By better understanding the correlates of school funding adequacy, we can take a major step towards implementing policy that targets the areas and communities most affected by this issue.

For this purpose, we focus our analysis on two key research questions. First, we investigate how socioeconomic factors impact average life expectancy. **We seek to answer the question, "What is the impact of residential segregation, access to healthy foods, and health insurance coverage on a county's average life expectancy?".** This analysis includes a comparison across different regions to identify variations in the relationship between residential segregation and life expectancy. Second, **we seek to answer the question, "Do residential segregation and region impact the likelihood of a school being adequately funded?"**. This question allows us to assess how the racial makeup and geographic location—given that each region of the U.S. is characterized by differing political and cultural norms—affect a county's educational investment. The observational unit for this analysis is the U.S. county.

To obtain data on the aforementioned variables, we used three primary sources.

The first source was the County Health Rankings & Roadmaps program,^3^ associated with the University of Wisconsin Population Health Institute, which provided us with 2024 datasets for all U.S. states. This program consolidates the latest county-level measurements of various population health, economic, demographic, and social factors from the American Community Survey, the USDA Food Environment Atlas, Small Area Income and Poverty Estimates, and other reliable government sources into publicly available datasets every year. Our specific variables of interest were originally collected as part of:

-   National Center for Health Statistics - Natality & Mortality Files, 2019-2021 (average life expectancy)

-   American Community Survey 5-year estimates, 2018-2022 (residential segregation index & median household income)

-   USDA Food Environment Atlas, 2019 (percentage of population with limited access to healthy foods)

-   Small Area Health Insurance Estimates, 2021 (percentage of adults uninsured)

-   School Finance Indicators Database, 2021 (school funding adequacy)

The second source was IPUMS, which provided us with median poverty tax data for every county from 2018-2022.^4^

The third source was the U.S. Census Bureau’s Regions and Divisions of the United States, which maps every U.S. to one of four geographic regions, and also a division within the regions. We specifically included regions as a variable in this analysis. This data was loaded into R using a [publicly available GitHub repository](https://github.com/cphalpert/census-regions) titled *Census Regions*.^5^

## **Methods**

### Variable Selection

Variables of interest were selected *a priori* based on their perceived relevance to the two outcomes we wanted to better understand: average life expectancy and school funding adequacy. While nutrition and access to quality healthcare are inherently linked to general wellbeing, it is worthwhile to assess how influential the environmental and financial aspects of these factors are on our tangible outcome. Therefore, we selected two key predictors for our life expectancy model: the percentage of the population with limited access to healthy foods and the percentage of uninsured adults in a county. The former is defined by the US Department of Agriculture as "the percentage of the \[county\] population that is low-income and does not live close to a grocery store", where "close" is defined differently for rural and nonrural areas.^6^ The latter variable represents the percentage of adults in a county younger than 65 who lack health insurance. We also were interested in seeing how residential segregation would impact life expectancy among these predictors, given existing studies on the negative health effects of segregation, and whether this relationship would exhibit any variations across different regions. This is why we chose to include residential segregation index values as a predictor. The residential segregation index, used by the County Health Rankings & Roadmaps program, measures how evenly two groups, such as Black and white residents, are distributed across geographic areas, with values ranging from 0 (complete integration) to 100 (complete segregation).^7^

Past studies have established a correlation between income and life expectancy.^8^ Because we wanted to control for its effects and examine how insurance, healthy food access, segregation, and region influence life expectancy even within a given income bracket, we selected income as our final predictor for the first model.

School funding adequacy is defined by the County Health Rankings program as "the average dollar gap between actual per-pupil spending and the amount needed for students to achieve national average test scores, considering districts' varying equity-based needs."^9^ To explore how racial diversity interface with variations in school funding, we selected residential segregation to be a part of this regression. In addition, given that state policy has a large influence on school administration and the determination of school needs, and that states in the same region tend to have similar political leanings and cultural beliefs, we wanted to analyze how much of a relationship exists between region and actual school funding levels. Because a major source of public school funding is tax revenue, median property tax and median household income were also included in the school funding model as potential confounders. This allowed us to assess whether the effects of segregation and region on funding existed outside of the effects of known economic factors.

### Exploratory Data Analysis & Data Preprocessing

2024 datasets for all U.S. states were combined and merged with median property tax and region mapping data, resulting in a final dataset with 3144 county observations and 9 different variables (state, average life expectancy, residential segregation index, % limited access to healthy foods, % uninsured adults, school funding adequacy, median household income, median property tax, and region). Because school funding adequacy was given as a numerical value, and we wanted to focus on better understanding the dichotomy between school funding adequacy and inadequacy, this variable was converted into a binary categorical variable (where the value was set to 1 if the numerical value was positive, indicating the counties’ schools were adequately funded, and set to 0 if the numerical value was negative, indicating underfunding).

Exploratory data analysis for the two outcome variables was conducted by 1) calculating descriptive statistics, 2) creating scatterplots of each individual predictor against the outcome (for life expectancy), and 3) creating pairwise plots between variables (for school funding). During the initial stages of analysis, we found that ⅓ of our observations were missing residential segregation index data, and that certain states were missing nearly all of their segregation values. Because of this, we decided to drop all states missing over 50% of their values for any variable. This led to our analysis being limited to 33 states, and excluding the following: Alaska, Colorado, Idaho, Iowa, Kansas, Minnesota, Montana, Nebraska, North Dakota, South Dakota, Utah, Wyoming, and Vermont. We were left with 2379 total observations, and the missing percentage for residential segregation dropped to 22%.

After this, the remaining missing values in all variables were imputed using the mice package and predictive mean matching to generate one complete dataset. The original school funding adequacy numerical variable was first imputed, and then converted into binary values for subsequent analysis.

### Model Fitting & Assessment

For our first research question, a multiple linear regression model was fit, regressing average life expectancy on residential segregation index, percentage with limited access to healthy foods, percentage uninsured adults, region, and median household income. Linear regression assumptions were assessed using diagnostic plots, especially residual vs. fitted and quantile-quantile plots. The adjusted R-squared value was used to evaluate the fit of the model, and the effect of region as an interaction term with residential segregation was evaluated using nested F tests. We also calculated VIF to check for multicollinearity and Cook’s distance to identify influential points.

For our second research question, a binary logistic regression model was fit with school funding adequacy (categorical) as the outcome and residential segregation index, region, median property tax, and median household income as predictors. The model was assessed using a confusion matrix, an ROC curve, and a comparison of deviance between the full model and a reduced model that excluded region as a predictor. VIF and Cook's distance were again used to evaluate multicollinearity and influential points.

## Results

### Data Overview

Table 1 displays summary statistics for the predictors in the second model, stratified by school funding adequacy. Table A1 in the Appendix provides summary statistics for all variables, and shows that both the mean and median for life expectancy are approximately 75 years. School funding adequacy counts show that around 66% of the counties in our analysis had underfunded schools. Additionally, the number of counties per region indicate an imbalance between categories, with the South region being over represented (holding 60% of all counties), and the East and Northeast holding around 9% of the counties each.

```{r, echo=FALSE, include=FALSE}
completed_data |> group_by(School.Funding.Cat) |> summarise(mean(Median.Household.Income), sd(Median.Household.Income), mean(Median.Prop.Tax), sd(Median.Prop.Tax), mean(Residential.Segregation.Index), sd(Residential.Segregation.Index))
```

```{r, echo=FALSE, include=FALSE}
table(completed_data$School.Funding.Cat, completed_data$Region_fac)
```

\newpage

*Table 1:* *School Funding Adequacy Summary Statistics*

| Variable                                 | **Inadequate School Funding** | **Adequate School Funding** |
|------------------------------|----------------------|--------------------|
| Residential Segregation Index: Mean (SD) | 46 (17)                       | 56 (14)                     |
| Region-West: Count (%)                   | 114 (7.3%)                    | 89 (10.8%)                  |
| Region-Northeast: Count (%)              | 3 (0.2%)                      | 200 (24.3%)                 |
| Region-Midwest: Count (%)                | 267 (17.1%)                   | 285 (34.7%)                 |
| Region-South: Count (%)                  | 1173 (75.3%)                  | 248 (30.2%)                 |
| Median Property Tax, USD: Mean (SD)      | 1119 (763)                    | 2282 (1648)                 |
| Median Household Income, USD: Mean (SD)  | 56513 (13200)                 | 72218 (18255)               |

```{r, echo=FALSE, include=FALSE}
# Plot for Q2 EDA
ggpairs(
  completed_data,
  columns = c("Residential.Segregation.Index", "Median.Household.Income", "Median.Prop.Tax"),
  aes(color = factor(School.Funding.Cat,
                     labels = c("Inadequate", "Adequate")))
) +
  labs(
    title = "Pairwise Plots of Predictors Colored by School Funding Adequacy",
    color = "School Funding Adequacy"
  ) +
  theme(
    plot.title = element_text(size = 15, hjust = 0.5),
    strip.text.x = element_text(size = 5),
    strip.text.y = element_text(size = 3)
  )

```

### Question 1

#### Model Results

Based on the multiple linear regression results, residential segregation, percentage of uninsured adults, region, and median household income all showed statistically significant effects on life expectancy. Specifically, for every \$10,000 increase in median household income, life expectancy increases by 1.25 years on average, holding other variables constant. In addition, the impact of region on life expectancy was substantial: life expectancy is on average 1.96 years higher in the West region compared to the South, 2.92 years higher in the Northeast compared to the South, and 1.92 years higher in the Midwest compared to the South, even when controlling for income, uninsured population, and racial diversity. This suggests, income as a factor withstanding, the differing quality of and access to healthcare and public services found in each region can have a significant impact on life expectancy.

Surprisingly, holding all other variables constant, for every 1 unit increase in percentage of uninsured adults in a county, a county's average life expectancy increases by 0.032 years. While this was statistically significant, many studies covering the relationship between an individual’s health and having insurance coverage have proven its positive correlation, suggesting our results may benefit from a closer look. In assessing our dataset, as seen in Appendix A1, we find that our dataset contains more observations for counties with low uninsured individuals, or equivalently, counties where most people are covered. As the timeframe of our dataset is 2018-2022, it is likely that the aforementioned observations require greater scrutiny of the significance of the impact of this variable on average life expectancy of a county.

The percentage of population with limited access to healthy foods predictor was not statistically significant in our model, suggesting that the effect of this environmental factor is minor or intertwined with economic variables, such as income, rather than being a standalone influence. In terms of residential segregation, for every 1 unit increase in a county's index value (corresponding to increased segregation), its average life expectancy increases by 0.0071 years in the South, increases by 0.0068 years in the West, decreases by 0.0055 years in the Northeast, and decreases by 0.0004 years in the Midwest, as seen in Figure 1. However, these regional differences are not statistically significant at the 0.05 significance level given the high p-values of all residential segregation-region interaction terms. We confirmed this result with a nested F-test, comparing a full model to a reduced model that excluded the interaction term between region and residential segregation. This did not show a statistically significant difference in RSS (*p*=0.624), indicating that the interaction between region and residential segregation does not add meaningful explanatory power to our model, Consequently, we did not find sufficient evidence to support regional variations in the relationship between segregation and life expectancy. Instead, based on residential segregation's coefficient estimate in the model, life expectancy does appear to increase overall with an increase in residential segregation, however minor this increase may be. One possible theory for this observation is that counties with more racial uniformity might experience less racial conflict and more social cohesion, especially for minority communities, but this should be further investigated in future studies.

*Figure 1: Segregation vs. Life Expectancy Across Different U.S. Regions*

```{r, echo=FALSE, fig.pos='H'}
library(ggplot2)
library(GGally)
library(patchwork)

ggplot(completed_data, aes(x = Residential.Segregation.Index, y = Life.Expectancy, color=Region_fac)) +
  geom_point(color='black') + geom_smooth(method="lm",se=F) + 
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Residential Segregation Index", y = "Average Life Expectancy (Years)", color='U.S. Region', title='Relationship Between Residential Segregation and Life Expectancy')+ 
  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
axis.title.x = element_text(size = 8),
axis.title.y = element_text(size = 8),
title=element_text(size=9))
```

\newpage

*Table 2: MLR Results For Average Life Expectancy Outcome*

```{r, echo= FALSE, warning = FALSE, fig.pos='H'}
# Running a regression with the imputated data (completed_data) to understand the impact of missing data

mlr.imp <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
    X..Limited.Access.to.Healthy.Foods +
    X..Uninsured.Adults + Median.Household.Income, data=completed_data)

renamed <- c(
    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
  "X..Limited.Access.to.Healthy.Foods" = "Percentage With Access to Healthy Foods",
  "X..Uninsured.Adults"= "Percentage of Uninsured Adults",
  "Region_facWest"="Region-West", 
  "Region_facNortheast"="Region-Northeast",
  "Region_facMidwest"="Region-Midwest" ,
  "Residential.Segregation.Index:Region_facWest" = "RSI*Region-West",
  "Residential.Segregation.Index:Region_facNortheast"= "RSI*Region-Northeast",
  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest",
  "Median Household Income" = "Median.Household.Income")

modelsummary(mlr.imp, fmt = fmt_significant(2),
             coef_rename = renamed, 
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA,
             )
```

```{r echo= FALSE, warning = FALSE, include=FALSE}
# Running a regression without the interaction term between residential segregation and region

mlr.imp.reduced <- lm(Life.Expectancy ~ Residential.Segregation.Index +
    X..Limited.Access.to.Healthy.Foods +
    X..Uninsured.Adults + Median.Household.Income +
    Region_fac, data=completed_data)

renamed <- c(
   "Residential Segregation Index (RSI)" = "Residential.Segregation.Index",
  "Percentage With Limited Access to Healthy Foods" = "X..Limited.Access.to.Healthy.Foods",
  "Percentage of Uninsured Adults"= "X..Uninsured.Adults",
  "Region-West" = "Region_facWest", 
  "Region-Northeast"= "Region_facNortheast",
  "Region-Midwest"= "Region_facMidwest",
  "Median Household Income" = "Median.Household.Income")

modelsummary(mlr.imp.reduced, fmt = fmt_significant(2),
             coef_rename = renamed, 
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA,
             )

```

```{r, include=FALSE, echo=FALSE}
##Nested F-Test to see if segregation-region interaction term makes a difference
anova(mlr.imp, mlr.imp.reduced)
```

#### Model Assessment

The adjusted R-squared value for the model was 0.6065, indicating that a majority of the variability in life expectancy was explained by the socioeconomic and racial predictors included in the model. In conducting diagnostics on the multilinear regression model, we first checked to see if the model met assumptions of linearity, independence, normality, and homoscedasticity. The residuals vs. fitted plot showed relatively equal variance and no discernible pattern, meaning that the linearity and homoscedasticity assumptions were reasonably met. In addition, because each observation in the dataset represented a unique county in the U.S., independence could be reasonably assumed. Assessing the Q-Q Plot for normality, the residual distribution was approximately normal, despite a slight divergence on both tails of the plot. Cook's distance was also assessed using the residuals vs. leverage plot; no points were found to be influential. There was a high leverage point (Mineral County, Nevada), and refitting the model without this point yielded no substantial changes in the coefficient estimates for statistically significant predictors; however, the *p*-value for residential segregation increased from 0.031 to 0.054. Finally, with a VIF \<2 for all predictors, there was little to no multicollinearity among the predictors.

```{r, echo=FALSE, include=FALSE}
completed_data[1180,]
completed_data_lev <- completed_data[-1180,]
```

```{r, echo=FALSE, include=FALSE}
#looking for outliers 
outlier_points <- outlierTest(mlr.imp)
print(outlier_points)
outlier_indices <- as.numeric(rownames(outlier_points))  # Extract row indices
print(outlier_indices)  # View the indices to confirm
```

```{r, echo=FALSE, include=FALSE}
#Variance Inflaction Factor (VIF)

#vif(mlr.imp)
vif(mlr.imp.reduced)
```

```{r, echo=FALSE, include=FALSE}
#Generate 4 Diagnostic Plots at once
par(mfrow = c(2, 2), mar = c(4,4,2,2))
plot(mlr.imp)
```

```{r, echo= FALSE, warning = FALSE, include=FALSE}
# Running a regression with obs 1180 removed

mlr.lev <- lm(Life.Expectancy ~ Residential.Segregation.Index * Region_fac +
    X..Limited.Access.to.Healthy.Foods +
    X..Uninsured.Adults + Median.Household.Income, data=completed_data_lev)

renamed <- c(
    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
  "X..Limited.Access.to.Healthy.Foods" = "Percentage With Access to Healthy Foods",
  "X..Uninsured.Adults"= "Percentage of Uninsured Adults",
  "Region_facWest"="Region-West", 
  "Region_facNortheast"="Region-Northeast",
  "Region_facMidwest"="Region-Midwest" ,
  "Residential.Segregation.Index:Region_facWest" = "RSI*Region-West",
  "Residential.Segregation.Index:Region_facNortheast"= "RSI*Region-Northeast",
  "Residential.Segregation.Index:Region_facMidwest"="RSI*Region-Midwest",
  "Median Household Income" = "Median.Household.Income")

modelsummary(mlr.lev, fmt = fmt_significant(2),
             coef_rename = renamed, 
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA,
             )
```

```{r, echo=FALSE, include=FALSE}
plot(mlr.lev)
```

## Question 2

#### Model Results

Table 3 below displays the odds ratio results from the logistic regression model for school funding adequacy, with the South as the baseline category for the categorical region variable. While all estimates are statistically significant at the 0.05 significance level, exponentiated regression coefficients from the logistic model demonstrate that a county's median household income and median property tax have almost no impact on the odds of its schools being adequately funded. Our results demonstrate that a one unit increase to the residential segregation index increases the odds of school funding adequacy by 0.02 times, allowing us to conclude that residential segregation has a minimal effect on the odds of the school being adequately funded. Figure 2 demonstrates an interesting trend—each region shows a different relationship between the residential segregation distribution of adequately and inadequately funded schools, pointing to a potential interaction between region and segregation.

Region has a substantial impact on the odds of school funding adequacy. Being located in the Northeast or the Midwest has the greatest impact on how well a school is funded, with the odds of adequate school funding being over 200 times higher in a Northeast county compared to a Southern county, and the odds of adequate school funding being over 3 times higher in a Midwest county compared to a Southern county, holding all other variables constant. Such disparities across the country are caused by a number of factors, including: (1) capacity - how well off a state is based on its economy and resources, and (2) effort - the state's willingness to provide funding for education. Wealthier states with a high fiscal capacity, (typically those in the Northeast), have more funding available to spend on education than states with more limited resources (typically those in the South and the West).

In comparing our full model (with predictor variables for residential segregation, median household income, median property tax, and region) to a model including all of these variables except for region, we found that there is statistical significance in the deviance between the residuals of the full model and reduced. As such, we can conclude that including region as a predictor refines and creates a better fit for our model.

```{r, echo=FALSE, include=FALSE}
## Checking to see if the imputated data has a large effect on the outcome

# Logistic Regression without Imputated Data
# Outcome = School.Funding.Cat
# Predictors = residential segregation, median.household.income, median.prop.tax

lmod <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac,
              data=combined,
              family="binomial")

#log-odds scale
modelsummary(lmod,
             fmt = fmt_significant(7),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA)

modelsummary(lmod,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)

```

```{r, include=FALSE, echo=FALSE}
imp_mods <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax + Region_fac, data=completed_data, family="binomial")

#log-odds scale
modelsummary(imp_mods,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = FALSE,
             gof_map=NA)
```

*Figure 2: Relationship Between Residential Segregation, Region, & School Funding Adequacy*

```{r}
ggplot(completed_data, aes(x=Region_fac, y = Residential.Segregation.Index, color=factor(School.Funding.Cat))) +
  geom_boxplot() +
  labs(x = "U.S. Region", y='Residential Segregation Index', color='School Funding Adequacy', title='Residential Segregation Stratified by Region & School Funding') +
  scale_color_hue(labels = c("Inadequate", "Adequate"))+ 
  theme(plot.margin = margin(t = 20, r = 30, b = 20, l = 10),
axis.title.x = element_text(size = 8),
axis.title.y = element_text(size = 8),
title=element_text(size=9))
```

*Table 3: Logistic Regression Results For School Funding Adequacy Outcome*

```{r, echo=FALSE}

renamed <- c(
    "Residential.Segregation.Index" = "Residential Segregation Index (RSI)",
    "Median.Household.Income"="Median Household Income, USD",
    "Median.Prop.Tax" = "Median Property Tax, USD",
  "Region_facWest"="Region-West", 
  "Region_facNortheast"="Region-Northeast",
  "Region_facMidwest"="Region-Midwest"
)
modelsummary(imp_mods,
             fmt = fmt_significant(2),
             shape = term ~ model + statistic,
             coef_rename = renamed,
             statistic = c("std.error","conf.int","p.value"),
             exponentiate = TRUE,
             gof_map=NA)
```

#### Model Assessment

To assess logistic regression assumptions, we first plotted each numerical predictor against the log odds of school funding adequacy. All plots showed a mostly linear relationship, indicating that the linearity assumption was largely met; interestingly, different linear trends were observed within each plot based on region, further supporting potential interaction effects between region and the three numerical predictors. GVIF for all predictors ranged between 1 and 3, suggesting that there could be moderate multicollinearity between predictors, but not enough to warrant a reconsideration of included variables. Cook's distance was also less than 0.05 for all points, meaning that there were no highly influential observations in the model. Evaluating our model, we found that the logistic regression model demonstrated a strong overall performance with an accuracy rate of 80.33%. The model had a high sensitivity of 91%, indicating that it accurately predicted 91% of actual positive cases of adequate school funding when evaluated on our dataset. However, it had a 60% specificity rate, highlighting it correctly identified cases of inadequate school funding only 60% of the time. Overall, looking at the ROC curve below, the AUC is 0.858, indicating a moderately strong discriminate ability.

```{r, include=FALSE, echo=FALSE}
plot(completed_data$Residential.Segregation.Index, predict(imp_mods), col = completed_data$Region_fac)
plot(completed_data$Median.Household.Income, predict(imp_mods), col = completed_data$Region_fac)
plot(completed_data$Median.Prop.Tax, predict(imp_mods), col = completed_data$Region_fac)
```

```{r, include=FALSE, echo=FALSE}
#Cooks Distance 
cooksd <- cooks.distance(imp_mods)

# Plot Cook's Distance
plot(cooksd, main = "Cook's Distance", ylab = "Cook's Distance")
abline(h = 4/(nrow(completed_data) - length(coef(imp_mods))), col = "red")  # Threshold line
4/(nrow(data) - length(coef(imp_mods)))
```

```{r, echo=FALSE, include=FALSE}
vif(imp_mods)
plot(imp_mods)
```

```{r, echo=FALSE, include=FALSE}
# Creating the confusion matrix

# Make predictions on the test set
test_probs <- predict(imp_mods, newdata = completed_data, type = "response")
test_pred <- ifelse(test_probs > 0.5, 1, 0)  # Classify based on threshold 0.5
```

```{r, echo=FALSE, include=FALSE}
# Create a confusion matrix
conf_matrix <- confusionMatrix(factor(test_pred), factor(completed_data$School.Funding.Cat))
#print(conf_matrix)

# Format and display confusion matrix as a Markdown table
pander(conf_matrix, caption = "Confusion Matrix")

```

```{r, echo=FALSE, include=FALSE}
#Creating the AUC curves
roc_result <- roc(
  completed_data$School.Funding.Cat,
  predict(imp_mods, type = "response"),
  legacy.axes = TRUE,
  plot = TRUE,
  print.thres = 0.5,
  print.auc = FALSE  # Disable automatic AUC
)

# Add AUC manually
text(
  x = 1.2,  # Adjust X-coordinate
  y = 1,  # Adjust Y-coordinate
  labels = sprintf("AUC = %.3f", auc(roc_result)),  # Add AUC value
  cex = 1,  # Label size
  col = "red"  # Label color
)
```

```{r, include=FALSE, echo=FALSE}
imp_mods_red <- glm(School.Funding.Cat ~ Residential.Segregation.Index + Median.Household.Income + Median.Prop.Tax, data=completed_data, family="binomial")

anova(imp_mods_red, imp_mods, test="Chisq")
```

## **Conclusion**

In Research Question 1, we used a multiple linear regression (MLR) model to assess how socioeconomic factors affect average life expectancy. The analysis shows that life expectancy is influenced by segregation, uninsured rates, household income, and regional differences, with the Northeast having the highest life expectancy, followed by the Midwest, West, and South. In Research Question 2, we used a logistic regression model to examine how residential segregation, property taxes, and median household income influence school funding adequacy. The results show that region has the greatest impact, with schools in the Northeast and Midwest being significantly more likely to receive adequate funding than those in the South. While residential segregation, property taxes, and household income have statistically significant effects, their influence is minimal compared to regional differences.

In our analysis, we observed the significant impact of the regional location of a county on both its average life expectancy (a health indicator) and school funding adequacy (an educational indicator)--two key components of societal and individual well-being. Life expectancy is highest in the Northeast, followed by the Midwest, West, and South, while school funding adequacy is significantly greater in the Northeast and Midwest compared to the South. Our findings highlight substantial differences in both outcomes between regions, which are often caused by different economic capabilities and different allocations of resources among countries. This study provides empirical evidence that underscores the importance for equitable social policies and targeted investments to address regional disparities and promote well-being across all communities.

Our analysis faced limitations due to missing data, which restricted the dataset to 33 states and reduced the generalizability of our findings. Additionally, some variables were drawn from different years between 2018 and 2022, requiring us to assume that trends remained consistent over a 1–5 year period and that time was an insignificant factor; however, this may not actually be the case.

Future studies could look at additional interaction effects with region and incorporate more social variables to enhance the analysis. For example, integrating a dataset on employment rates and school performance could offer deeper insights into how socioeconomic factors shape life expectancy and educational funding, complementing existing variables in our models. Another direction for future work is to delve deeper into understanding the surprising findings regarding the effects of uninsured adult percentage and residential segregation on life expectancy. Expanding the dataset to include a temporal scope, and finding more complete sources of data on variables of interest so that no states have to be excluded from the analysis, would also contribute to a more comprehensive understanding.

## **Appendix**

*Table A1: Summary Statistics For All Variables*

```{r, include=FALSE, echo=FALSE}
renamed_data <- completed_data %>%
  rename(
    "Residential Segregation" = Residential.Segregation.Index,
    "Median Household Income" = Median.Household.Income,
    "Median Property Tax" = Median.Prop.Tax,
    "School Funding Adequacy" = School.Funding.Adequacy, 
    "Life Expectancy" = Life.Expectancy,
    "% Limited Access, Healthy Foods" = X..Limited.Access.to.Healthy.Foods,
    "% Uninsured Adults" = X..Uninsured.Adults,
    "Region" = Region_fac,
    "School Funding Binary" =School.Funding.Cat)

skim_table <- datasummary_skim(
  renamed_data,
  title = "Summary Statistics")

```

## ![](images/clipboard-2912764107.png)

## **References**

1.  National Academies of Sciences, Engineering, and Medicine. (2017). *Communities in action: Pathways to health equity.* Washington, DC: The National Academies Press. Retrieved from <https://www.ncbi.nlm.nih.gov/books/NBK425845/>

2.  Stanford Center for Opportunity Policy in Education. (2013.). How to close the opportunity gap: Key policy recommendations. Retrieved from <https://edpolicy.stanford.edu/sites/default/files/Opp%20Gap%20Policy%20Recommendations.pdf>

3.  County Health Rankings & Roadmaps. (n.d.). *About us.* Retrieved December 8, 2024, from <https://www.countyhealthrankings.org/about-us>

4.  Steven Ruggles, Sarah Flood, Matthew Sobek, Daniel Backman, Annie Chen, Grace Cooper, Stephanie Richards, Renae Rodgers, and Megan Schouweiler. IPUMS USA: Version 15.0 \[dataset\]. Minneapolis, MN: IPUMS, 2024. Retrieved from <https://doi.org/10.18128/D010.V15.0>

5.  Halpert, C. (2014). Census regions \[GitHub repository\]. GitHub. <https://github.com/cphalpert/census-regions>

6.  USDA. (2022). *Documentation*. Economic Research Service. https://www.ers.usda.gov/data-products/food-access-research-atlas/documentation/

7.  County Health Rankings & Roadmaps. (2024). *Residential segregation - Black/White.* University of Wisconsin Population Health Institute. Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/family-and-social-support/residential-segregation-blackwhite?utm_source=chatgpt.com&year=2024>

8.  Chetty, R., Stepner, M., Abraham, S., Lin, S., Scuderi, B., Turner, N., Bergeron, A., & Cutler, D. (2016). The Association Between Income and Life Expectancy in the United States, 2001-2014. *JAMA*, *315*(16), 1750–1766. https://doi.org/10.1001/jama.2016.4226

9.  County Health Rankings & Roadmaps. (2024). *School funding adequacy.* University of Wisconsin Population Health Institute. Retrieved from <https://www.countyhealthrankings.org/health-data/health-factors/social-economic-factors/education/school-funding-adequacy?utm_source=chatgpt.com&year=2024>